---
title: "FE_PROJECT"
author: "Gautam Prasad"
date: "2025-07-03"
output: html_document
---

# Library
```{r echo=TRUE, message=FALSE, warning=FALSE}
library(quantmod)
library(tseries)
library(FinTS)
library(fracdiff)
library(lmtest)
library(forecast)
library(ggplot2)
library(reshape2)
library(scales)
library(rugarch)
library(PerformanceAnalytics)
library(urca)
library(zoo)
library(xts)
```

# Data (Stock - Reliance & Index - Nifty 50)
``` {r}
getSymbols(c("^NSEI", "RELIANCE.NS"), src = "yahoo", from = "2021-10-01", to = "2025-01-01", periodicity = "d")

merged_prices <- merge(Cl(NSEI), Cl(RELIANCE.NS))  
colnames(merged_prices) <- c("Nifty50", "Reliance")

#Calculate Daily Log Returns (in percent)
log_returns <- na.omit(100 * diff(log(merged_prices)))
stock <- log_returns$Reliance
index <- log_returns$Nifty50

```

## Plots (log return, Acf & Pacf)
```{r}
par(mfrow = c(2, 1))
plot(stock, main = "Reliance Log Returns", col = "black", ylab = "Log Return", xlab = "Date", type = "l")
plot(index, main = "Nifty 50 Log Returns", col = "skyblue", ylab = "Log Return", xlab = "Date", type = "l")

par(mfrow = c(2, 2))
asset_name = "Reliance"
Acf(stock, main = paste("Reliance - ", "ACF"))
Pacf(stock, main = paste("Reliance - ", "PACF"))

asset_name = "Nifty 50"
Acf(index, main = paste("Nifty 50 - ", "ACF"))
Pacf(index, main = paste("Nifty 50 - ", "PACF"))

```

##  Test (Normality,Unit Root Test, ARCH Effect Test, Long Memory Test, signbias_test)

# Normality Test
```{r}
shapiro.test(coredata(stock)) 
# Data is not normally distributed 

shapiro.test(coredata(index))
# Data is not normally distributed

cat("Both Stock and Index are ** Non-Normal Distributed **")


```

##  Unit Root ADF
```{r}
adf.test(stock) 
# staionary data
adf.test(index) 
# staionary data

cat("Both Stock and Index are ** Stationary **")
```

## Long memory test

```{r}
fracdiff(stock)
# d = 4.58e-05 (≈ 0) => does not exhibit long memory

fracdiff(index)
# d = 4.583013e-05 (≈ 0) => does not exhibit long memory

cat("Both Stock and Index have ** Short memory **")


```

##  ARCH test
```{r}
ArchTest(stock, lags = 10)
#  Arch-Effect is Present  

ArchTest(index, lags = 10)
# Arch-Effect is Present  

cat("Both Stock and Index have ** significant ARCH effects **")



```

#  Sign Bias test
```{r}

garch_spec <- ugarchspec(
  variance.model = list(model = "sGARCH", garchOrder = c(1,1)),
  mean.model     = list(armaOrder = c(1,1), include.mean = TRUE),
  distribution.model = "norm"
)

garch_fit_stock <- ugarchfit(spec = garch_spec, data = stock)
garch_fit_index <- ugarchfit(spec = garch_spec, data = index)


signbias(garch_fit_stock)
signbias(garch_fit_index)

cat("Both Stock and Index have ** No significant asymmetry  **")


```

#  ARMA Model - stock (Reliance)
```{r}

# ARMA (0,0)
arma00 <- arima(stock, order = c(0,0,0)) # log likelihood = -1437.64,  aic = 2879.28
summary(arma00)

# ARMA(1,0)
arma10 <- arima(stock, order = c(1,0,0)) # log likelihood = -1437.53,  aic = 2881.07
summary(arma10)

# ARMA(0,1)
arma01 <- arima(stock, order = c(0,0,1)) # log likelihood = -1437.54,  aic = 2881.07
summary(arma01)

# ARMA(1,1)
arma11 <- arima(stock, order = c(1,0,1)) # log likelihood = -1436.76,  aic = 2881.53
summary(arma11)

# ARMA(2,0)
arma20 <- arima(stock, order = c(2,0,0)) #  log likelihood = -1437.47,  aic = 2882.94
summary(arma20)

# ARMA(0,2)
arma02 <- arima(stock, order = c(0,0,2)) #  log likelihood = -1437.5,  aic = 2882.99
summary(arma02)

# ARMA(2,2)
arma22 <- arima(stock, order = c(2,0,2)) # log likelihood = -1432.85,  aic = 2877.71
summary(arma22)

#ARMA(3,0)
arma30 <- arima(stock, order = c(3,0,0)) # log likelihood = -1434.36,  aic = 2878.72
summary(arma30)

#ARMA(0,3)
arma03 <- arima(stock, order = c(0,0,3)) # log likelihood = -1434.68,  aic = 2879.35
summary(arma03)

# ARMA(3,3)
arma33 <- arima(stock, order = c(3,0,3)) # log likelihood = -1431.77,  aic = 2879.54
summary(arma33)


```


##  Best Model - Stock (reliance) 
```{r } 
cat("Best Model is ---->  ** arma(2,2) ** ")

# Among all the model, 
#(2,2) has lowest AIC with aic(2,2) = 2877.71,
# Also, coefficient ar1,ar2 and ma1, ma2 is statstically significant

arma_residuals <- residuals(arma22)

#Heteroskedasticity
ArchTest(arma_residuals, lags = 10)
# Presence of ARCH effects => consider GARCH models

# White noise
Box.test(arma_residuals, lag = 10, type = "Ljung-Box") # p-value = 0.3489
# Residuals are not autocorrelated and behave like white noise

# Normality test
shapiro.test(arma_residuals) 
# Residuals are not normally distributed (presence of tails, skewness, or kurtosis)

# ACF
asset_name = "(Residuals - Reliance) - "
acf(arma_residuals, main = paste(asset_name, "ACF"))

```


##  ARMA Model - Index (Nifty 500)
```{r}

# ARMA (0,0)
Arma00 <- arima(index, order = c(0,0,0)) # log likelihood = -1047.84,  aic = 2099.67
summary(Arma00)

# ARMA(1,0)
Arma10 <- arima(index, order = c(1,0,0)) # log likelihood = -1047.72,  aic = 2101.44
summary(Arma10)

# ARMA(0,1)
Arma01 <- arima(index, order = c(0,0,1)) # log likelihood = -1047.72,  aic = 2101.44
summary(Arma01)

# ARMA(1,1)
Arma11 <- arima(index, order = c(1,0,1)) #  log likelihood = -1047.25,  aic = 2102.51
summary(Arma11)

# ARMA(2,0)
Arma20 <- arima(index, order = c(2,0,0)) # log likelihood = -1047.69,  aic = 2103.39
summary(Arma20)

# ARMA(0,2)
Arma02 <- arima(index, order = c(0,0,2)) # log likelihood = -1047.7,  aic = 2103.4
summary(Arma02)

# ARMA(2,2)
Arma22 <- arima(index, order = c(2,0,2)) # log likelihood = -1046.07,  aic = 2104.14
summary(Arma22)

```


##  Best ARMA Model - Index (Nifty 50) 

```{r}

cat("Best Model is ---->  ** Arma(0,0) ** ")

# arma(0,0) has the lowest AIC = 2099.67

Arma_residuals <- residuals(Arma00)

#Heteroskedasticity
ArchTest(Arma_residuals, lags = 10)
# Presence of ARCH effects => consider GARCH models

# White noise
Box.test(Arma_residuals, lag = 10, type = "Ljung-Box") # p-value = 0.859
# Residuals are not autocorrelated and behave like white noise

# Normality test
shapiro.test(Arma_residuals) 
# Residuals are not normally distributed (presence of tails, skewness, or kurtosis)

# ACF
asset_name = "(Residuals - Nifty 50) - "
acf(Arma_residuals, main = paste(asset_name, "ACF"))



```

##  Garch Model - Stock (Reliance) With arma(2,2)

```{r}
# Mean model spec (ARMA(2,2) with mean)
mean_ <- list(armaOrder = c(2,2), include.mean = TRUE)

# GARCH(1,1)
spec_garch <- ugarchspec(
  variance.model = list(model = "sGARCH", garchOrder = c(1,1)),
  mean.model = mean_,
  distribution.model = "std"
)
fit_garch <- ugarchfit(spec_garch, data = stock)
show(fit_garch)


# GJR-GARCH(1,1)
spec_gjr <- ugarchspec(
  variance.model = list(model = "gjrGARCH", garchOrder = c(1,1)),
  mean.model = mean_,
  distribution.model = "std"
)
fit_gjr <- ugarchfit(spec_gjr, data = stock)
show(fit_gjr)

# EGARCH(1,1)
spec_egarch <- ugarchspec(
  variance.model = list(model = "eGARCH", garchOrder = c(1,1)),
  mean.model = mean_,
  distribution.model = "std"
)
fit_egarch <- ugarchfit(spec_egarch, data = stock)
show(fit_egarch)

# IGARCH(1,1)
spec_igarch <- ugarchspec(
  variance.model = list(model = "iGARCH", garchOrder = c(1,1)),
  mean.model = mean_,
  distribution.model = "std"
)
fit_igarch <- ugarchfit(spec_igarch, data = stock)
show(fit_igarch)

# APARCH(1,1)
spec_aparch <- ugarchspec(
  variance.model = list(model = "apARCH", garchOrder = c(1,1)),
  mean.model = mean_,
  distribution.model = "std"
)
fit_aparch <- ugarchfit(spec_aparch, data = stock)
show(fit_aparch)

# GARCH-in-Mean(1,1)
mean_garchm <- list(armaOrder = c(2,2), include.mean = TRUE, archm = TRUE)
spec_garchm <- ugarchspec(
  variance.model = list(model = "sGARCH", garchOrder = c(1,1)),
  mean.model = mean_garchm,
  distribution.model = "std"
)
fit_garchm <- ugarchfit(spec_garchm, data = stock)
show(fit_garchm)

# Compare Information Criteria 

infocriteria(fit_garch)
infocriteria(fit_gjr)
infocriteria(fit_egarch)
infocriteria(fit_igarch)
infocriteria(fit_aparch)
infocriteria(fit_garchm)

```


##  Best Garch Model - Stock (Reliance) 
```{r}

# GARCH(1,1):      Akaike (AIC): 3.465035, Bayes (BIC): 3.517685  -> Lowest AIC
# GJR-GARCH(1,1):  Akaike (AIC): 3.477226, Bayes (BIC): 3.535726
# EGARCH(1,1):     Akaike (AIC): 3.476196, Bayes (BIC): 3.534697
# IGARCH(1,1):     Akaike (AIC): 3.475146, Bayes (BIC): 3.521946
# APARCH(1,1):     Akaike (AIC): 3.467837, Bayes (BIC): 3.532187
# GARCH-M(1,1):    Akaike (AIC): 3.477146, Bayes (BIC): 3.535646 

cat("Best Model is ---->  **  garch(1,1)** ")
# Significant: (AR1, AR2, MA1, MA2, omega, alpha1, beta1, shape)
 
garch_residuals <- residuals(fit_garch, standardize = TRUE)

#Heteroskedasticity
ArchTest(garch_residuals, lags = 10) 
# p-value = 0.08495 > 0.05 => No presence of ARCH effects

# White noise
Box.test(garch_residuals, lag = 10, type = "Ljung-Box") 
# p-value = 0.3488 => Residuals are white noise

# ACF
asset_name = "(Garch Residuals - Reliance) - "
acf(garch_residuals, main = paste(asset_name, "ACF"))
# no significant autocorrelation 

```

##  Garch Model - Index (Nifty 50) With arma(0,0)
```{r}
# Mean model spec (ARMA(0,0) with mean)
Mean_ <- list(armaOrder = c(0,0), include.mean = TRUE)

# GARCH(1,1)
Spec_garch <- ugarchspec(
  variance.model = list(model = "sGARCH", garchOrder = c(1,1)),
  mean.model = Mean_,
  distribution.model = "std"
)
Fit_garch <- ugarchfit(Spec_garch, data = index)
show(Fit_garch)


# GJR-GARCH(1,1)
Spec_gjr <- ugarchspec(
  variance.model = list(model = "gjrGARCH", garchOrder = c(1,1)),
  mean.model = Mean_,
  distribution.model = "std"
)
Fit_gjr <- ugarchfit(Spec_gjr, data = index)
show(Fit_gjr)

# EGARCH(1,1)
Spec_egarch <- ugarchspec(
  variance.model = list(model = "eGARCH", garchOrder = c(1,1)),
  mean.model = Mean_,
  distribution.model = "std"
)
Fit_egarch <- ugarchfit(Spec_egarch, data = index)
show(Fit_egarch)

# EGARCH(2,1)
Spec_egarch21 <- ugarchspec(
  variance.model = list(model = "eGARCH", garchOrder = c(2,1)),
  mean.model = Mean_,
  distribution.model = "std"
)
Fit_egarch21 <- ugarchfit(Spec_egarch21, data = index)
show(Fit_egarch21)

# EGARCH(1,2)
Spec_egarch12 <- ugarchspec(
  variance.model = list(model = "eGARCH", garchOrder = c(1,2)),
  mean.model = Mean_,
  distribution.model = "std"
)
Fit_egarch12 <- ugarchfit(Spec_egarch12, data = index)
show(Fit_egarch12)



# IGARCH(1,1)
Spec_igarch <- ugarchspec(
  variance.model = list(model = "iGARCH", garchOrder = c(1,1)),
  mean.model = Mean_,
  distribution.model = "std"
)
Fit_igarch <- ugarchfit(Spec_igarch, data = index)
show(Fit_igarch)

# APARCH(1,1)
Spec_aparch <- ugarchspec(
  variance.model = list(model = "apARCH", garchOrder = c(1,1)),
  mean.model = Mean_,
  distribution.model = "std"
)
Fit_aparch <- ugarchfit(Spec_aparch, data = index)
show(Fit_aparch)

# APARCH(2,1)
Spec_aparch21 <- ugarchspec(
  variance.model = list(model = "apARCH", garchOrder = c(2,1)),
  mean.model = Mean_,
  distribution.model = "std"
)
Fit_aparch21 <- ugarchfit(Spec_aparch21, data = index)
show(Fit_aparch21)

# APARCH(1,2)
Spec_aparch12 <- ugarchspec(
  variance.model = list(model = "apARCH", garchOrder = c(1,2)),
  mean.model = Mean_,
  distribution.model = "std"
)
Fit_aparch12 <- ugarchfit(Spec_aparch12, data = index)
show(Fit_aparch12)

# APARCH(2,2)
Spec_aparch22 <- ugarchspec(
  variance.model = list(model = "apARCH", garchOrder = c(2,2)),
  mean.model = Mean_,
  distribution.model = "std"
)
Fit_aparch22 <- ugarchfit(Spec_aparch22, data = index)
show(Fit_aparch22)

# GARCH-in-Mean(1,1)
Mean_garchm <- list(armaOrder = c(0,0), include.mean = TRUE, archm = TRUE)
Spec_garchm <- ugarchspec(
  variance.model = list(model = "sGARCH", garchOrder = c(1,1)),
  mean.model = Mean_garchm,
  distribution.model = "std"
)
Fit_garchm <- ugarchfit(Spec_garchm, data = index)
show(Fit_garchm)

# Compare Information Criteria 

infocriteria(Fit_garch)
infocriteria(Fit_gjr)
infocriteria(Fit_egarch)
infocriteria(Fit_egarch21)
infocriteria(Fit_egarch12)
infocriteria(Fit_igarch)
infocriteria(Fit_aparch)
infocriteria(Fit_aparch21)
infocriteria(Fit_aparch12)
infocriteria(Fit_aparch22)
infocriteria(Fit_garchm)



```


##  Best Garch Model - Index (Nifty 50) with arma(0,0) 
```{r}
# GARCH(1,1):      Akaike (AIC): 2.459629, Bayes (BIC): 2.488879
# GJR-GARCH(1,1):  Akaike (AIC): 2.445246, Bayes (BIC): 2.480346
# EGARCH(1,1):     Akaike (AIC): 2.438112, Bayes (BIC): 2.473212
# EGARCH(2,1):     Akaike (AIC): 2.438112, Bayes (BIC): 2.473212
# EGARCH(1,2):     Akaike (AIC): 2.438112, Bayes (BIC): 2.473212
# IGARCH(1,1):     Akaike (AIC): 2.465607, Bayes (BIC): 2.489007
# APARCH(1,1):     Akaike (AIC): 2.437239, Bayes (BIC): 2.478189
# APARCH(2,1):     Akaike (AIC): 2.442094, Bayes (BIC): 2.494745
# APARCH(1,2):     Akaike (AIC): 2.438205, Bayes (BIC): 2.485006
# APARCH(2,2):     Akaike (AIC): 2.443199, Bayes (BIC): 2.501699
# GARCH-M(1,1):    Akaike (AIC): 2.461007, Bayes (BIC): 2.496107


cat("Best Model is ---->  **  aparch(1,1)** ")
# aparch (1,1) has lowest AIC, then egarch(1,1). 
# also, apARCH has all significant coefficients while eGARCH's ω (omega) is not significant.
# Tried Aparch(1,2), Aparch(2,1) - they have same AIC value and same case with egarach(1,2), egarach(2,1)
 
Garch_residuals <- residuals(Fit_aparch, standardize = TRUE)

#Heteroskedasticity
ArchTest(Garch_residuals, lags = 10) 
# p-value = 0.0009728 < 0.05 =>  presence of ARCH effects
# aparch(1,1) and egarch(1,1) are the two best model, but both have arch effect, even their high order.

# White noise
Box.test(Garch_residuals, lag = 10, type = "Ljung-Box") 
# p-value = 0.9329 => Residuals are white noise

# ACF
asset_name = "(Garch Residuals - Nifty 50) - "
acf(Garch_residuals, main = paste(asset_name, "ACF"))
# no significant autocorrelation 



```


# conditional variance - Stock (Reliance) with best Garch(1,1) - arma(2,2)
``` {r}
cond_var_garch      <- sigma(fit_garch)^2
cond_var_gjr        <- sigma(fit_gjr)^2
cond_var_egarch     <- sigma(fit_egarch)^2
cond_var_igarch     <- sigma(fit_igarch)^2
cond_var_aparch     <- sigma(fit_aparch)^2
cond_var_garchm     <- sigma(fit_garchm)^2

# Best Model
plot(cond_var_garch, type = "l", col = "blue", 
     main = "Conditional Variance - (Garch(1,1)+ arma(2,2))",
     xlab = "Time", ylab = "Conditional Variance")

plot(cond_var_garch, type = "l", col = "blue", 
     main = "Comparison of Conditional Variances ",
     xlab = "Time", ylab = "Conditional Variance",
     ylim = range(c(cond_var_garch, cond_var_gjr, cond_var_egarch, 
                    cond_var_igarch, cond_var_aparch, cond_var_garchm), na.rm = TRUE))
lines(cond_var_gjr, col = "red")
lines(cond_var_egarch, col = "green")
lines(cond_var_igarch, col = "purple")
lines(cond_var_aparch, col = "orange")
lines(cond_var_garchm, col = "brown")
legend("topright", 
       legend = c("GARCH", "GJR-GARCH", "EGARCH", "IGARCH", "APARCH", "GARCH-M"),
       col = c("blue", "red", "green", "purple", "orange", "brown"),
       lty = 1, cex = 0.8)


```

# conditional mean - Stock (Reliance) with best Garch(1,1) - arma(2,2)
``` {r}
cond_mean_garch <- fitted(fit_garch)

# Plot conditional mean
plot(cond_mean_garch, type = "l", col = "blue", lwd = 2,
     main = "Conditional Mean - Garch(1,1) +arma(0,0)",
     ylab = "Conditional Mean", xlab = "Time")

```

# conditional variance - Index (Nifty 50) with  aparch(1,1) + arma(0,0)
``` {r}

Cond_var_garch      <- sigma(Fit_garch)^2
Cond_var_gjr        <- sigma(Fit_gjr)^2
Cond_var_egarch     <- sigma(Fit_egarch)^2
Cond_var_igarch     <- sigma(Fit_igarch)^2
Cond_var_aparch     <- sigma(Fit_aparch)^2
Cond_var_garchm     <- sigma(Fit_garchm)^2

# Best Model
plot(Cond_var_garch, type = "l", col = "blue", 
     main = "Conditional Variance - aparch(1,1) + arma(0,0))",
     xlab = "Time", ylab = "Conditional Variance")

plot(Cond_var_garch, type = "l", col = "blue", 
     main = "Comparison of Conditional Variances ",
     xlab = "Time", ylab = "Conditional Variance",
     ylim = range(c(Cond_var_garch, Cond_var_gjr, Cond_var_egarch, 
                    Cond_var_igarch, Cond_var_aparch, Cond_var_garchm), na.rm = TRUE))
lines(Cond_var_gjr, col = "red")
lines(Cond_var_egarch, col = "green")
lines(Cond_var_igarch, col = "purple")
lines(Cond_var_aparch, col = "orange")
lines(Cond_var_garchm, col = "brown")
legend("topright", 
       legend = c("GARCH", "GJR-GARCH", "EGARCH", "IGARCH", "APARCH", "GARCH-M"),
       col = c("blue", "red", "green", "purple", "orange", "brown"),
       lty = 1, cex = 0.8)


```

# conditional mean - Index (Nifty 50) with aparch(1,1) + arma(0,0)
``` {r}
Cond_mean_garch <- fitted(Fit_aparch)

# Plot conditional mean
plot(Cond_mean_garch, type = "l", col = "blue", lwd = 2,
     main = "Conditional Mean - aparch(1,1) + arma(0,0)",
     ylab = "Conditional Mean", xlab = "Time")
```



# Forecast for Stock Reliance
``` {r}
# next 5 days Forecast
forecast <- ugarchforecast(fit_garch, n.ahead = 5)
sigma_forecast <- sigma(forecast)
show(sigma_forecast)

# Volatility is increasing slightly daily, which means model expects increase in risk (slightly) over next 5 days, won't #affect stability of market that much 

```

# Forecast for Index Nifty 50
``` {r}
# next 5 days Forecast
Forecast <- ugarchforecast(Fit_aparch, n.ahead = 5)
Sigma_forecast <- sigma(Forecast)
show(Sigma_forecast)

# Volatility is slowly decreasing slightly, model expects market to be less risky over the next 5 days, suggesting a calm # # period ahead.
```